---
title: "Data Cleaning and Basic Analyses"
author: "Thomas Gareau-Paquette"
date: "2024-03-07"
output: html_document
---

```{r}
library(haven)
library(tidyverse)
library(ggplot2)
library(ggtext)
library(plotly)
library(labelled)
library(survey)
library(scales)
library(fixest)
library(broom)
library(GGally)
library(readxl)
library(viridis)
library(ggrepel)
library(modelsummary)
library(kableExtra)
library(summarytools)
library(psych)
library(reshape2)
```

```{r}
ces2019 <- read_dta("ces2019_online/2019 Canadian Election Study - Online Survey v1.0.dta")
```



## Province

```{r}
ces2019 <- ces2019 %>%
    mutate(region = ifelse(pes19_province == 4 | pes19_province == 5 | pes19_province == 7 | pes19_province == 10, "Atlantic", NA),
           region = ifelse(pes19_province == 1 | pes19_province == 3 | pes19_province == 12, "Prairies", region),
           region = ifelse(pes19_province == 11, "Quebec", region),
           region = ifelse(pes19_province == 9, "Ontario", region),
           region = ifelse(pes19_province == 2, "BC", region)) %>%
    filter(!is.na(region))
```


# Controls

## education (4 categories)

```{r}
ces2019 <- ces2019 %>%
  mutate(education = case_when(
    cps19_education %in% 1:5 ~ "No Schooling/Completed High School",
    cps19_education == 6 ~ "Some CEGEP/Post-Secondary",
    cps19_education == 7 ~ "Completed CEGEP/Post-Secondary",
    cps19_education == 8 ~ "Some University",
    cps19_education == 9 ~ "Bachelor's Degree",
    cps19_education == 10 ~ "Master's Degree/Doctorate",
    cps19_education == 11 ~ "Master's Degree/Doctorate",
    TRUE ~ NA_character_  # assign NA to values that do not fit the specified categories
  )) %>%
  filter(!is.na(education)) %>%
    mutate(education = factor(education, levels = c("No Schooling/Completed High School", "Some CEGEP/Post-Secondary", "Completed CEGEP/Post-Secondary", "Some University", "Bachelor's Degree", "Master's Degree/Doctorate")))

```


## age

```{r}
ces2019 <- ces2019 %>%
  mutate(agegroup =  cps19_age,
         agegroup = ifelse(cps19_age <= 35, "18-35", agegroup),
         agegroup = ifelse(cps19_age >= 36 & cps19_age <= 45, "36-45", agegroup),
         agegroup = ifelse(cps19_age >= 46 & cps19_age <= 55, "46-55", agegroup),
         agegroup = ifelse(cps19_age >= 56 & cps19_age <= 65, "56-65", agegroup),
         agegroup = ifelse(cps19_age > 65, "66 +", agegroup)) %>%
    mutate(agegroup = factor(agegroup, levels = c("18-35", "36-45", "46-55", "56-65", "66 +")))
```



## Racial minority (ADDED 18/09/2023 `minovis`)

```{r}
# odd wording
ces2019 <- ces2019 %>%
  mutate(
    minovis = case_when(
      cps19_ethnicity_24 == 1 | cps19_ethnicity_26 == 1 | cps19_ethnicity_27 == 1 | 
      cps19_ethnicity_28 == 1 | cps19_ethnicity_29 == 1 | cps19_ethnicity_30 == 1 | 
      cps19_ethnicity_34 == 1 | cps19_ethnicity_35 == 1 | cps19_ethnicity_37 == 1 | 
      cps19_ethnicity_38 == 1 | cps19_ethnicity_39 == 1 | cps19_ethnicity_40 == 1 ~ 0,
      cps19_ethnicity_43 == 1 ~ NA_integer_,
      TRUE ~ 1
    )
  )
```


## Gender

```{r}
ces2019 <- ces2019 %>%
    filter(cps19_gender != 3) %>%
    mutate(gender = ifelse(cps19_gender == 1, 0, 1))
```


## Religiosity (3 categories)

```{r}
ces2019 <- ces2019 %>%
    mutate(has_religion = ifelse(cps19_religion == 1, "No Religion", "Has Religion"))
```

## Religious Frequency

```{r}
ces2019 <- ces2019 %>%
    mutate(religious_freq = ifelse(pes19_service_freq == 1, "Never/Almost never", NA),
           religious_freq = ifelse(pes19_service_freq == 2, "Never/Almost never", religious_freq),
           religious_freq = ifelse(pes19_service_freq == 3, "Never/Almost never", religious_freq),
           religious_freq = ifelse(pes19_service_freq == 4, "Once a month or more", religious_freq),
           religious_freq = ifelse(pes19_service_freq == 5, "Once a month or more", religious_freq),
           religious_freq = ifelse(pes19_service_freq == 6, "Once a month or more", religious_freq),
           religious_freq = ifelse(pes19_service_freq == 7, "Never/Almost never", religious_freq))
```




## Urban/Rural Context

```{r}
ces2019 <- ces2019 %>%
  mutate(urban_rural = pes19_rural_urban) %>%
  mutate(urban_rural = replace(urban_rural, urban_rural == 6, NA)) %>%
  mutate(urban_rural = as_factor(urban_rural))
```


## Language

```{r}
ces2019 <- ces2019 %>%
    mutate(language = ifelse(!is.na(cps19_language_68), "English", "Other"),
           language = ifelse(!is.na(cps19_language_69), "French", language)) %>%
    mutate(language = factor(language, levels = c("English", "French", "Other")))
```


# Attitudes toward redistribution

* pes19_govt_act_ineq: The government should take measures to reduce differences in income levels.	

* Strongly disagree (1)
* Somewhat disagree (2)
* Neither agree nor disagree (3)
* Somewhat agree (4)
* Strongly agree (5)
* Don't know/ Prefer not to answer (6)

```{r}
ces2019 <- ces2019 %>%
  mutate(govt_act_ineq = case_when(
    pes19_govt_act_ineq == 6 ~ NA_real_, 
    TRUE ~ (pes19_govt_act_ineq - 1) / 4 
  ))

```



* pes19_inequal: Is income inequality a big problem in Canada?

* Definitely yes (1)
* Probably yes (2)
* Not sure (3)
* Probably not (4)
* Definitely not (5)
* Don't know/ Prefer not to answer (6)

```{r}
ces2019 <- ces2019 %>%
  mutate(inequal = case_when(
    pes19_inequal == 6 ~ NA_real_, 
    TRUE ~ (5 - pes19_inequal) / 4
  ))
```


* pes19_stdofliving: The government should:

* See to it that everyone has a decent standard of living (1)
* Leave people to get ahead on their own (2)
* Don't know/ Prefer not to answer (3)
    
```{r}
# Directly recode the pes19_stdofliving variable
ces2019 <- ces2019 %>%
  mutate(stdofliving = case_when(
    pes19_stdofliving == 1 ~ 1,
    pes19_stdofliving == 2 ~ 0,
    pes19_stdofliving == 3 ~ 0.5  # Assign 0.5 to "Don't know/Prefer not to answer"
  ))
```

    
* pes19_privjobs: The government should leave it entirely to the private sector to create jobs.

* Strongly disagree (1)
* Somewhat disagree (2)
* Neither agree nor disagree (3)
* Somewhat agree (4)
* Strongly agree (5)
* Don't know/ Prefer not to answer (6)

```{r}
ces2019 <- ces2019 %>%
  mutate(privjobs = case_when(
    pes19_privjobs == 6 ~ NA_real_,
    TRUE ~ (5 - pes19_privjobs) / 4
  ))
```



* pes19_blame: People who don’t get ahead should blame themselves, not the system.

* Strongly disagree (1)
* Somewhat disagree (2)
* Neither agree nor disagree (3)
* Somewhat agree (4)
* Strongly agree (5)
* Don't know/ Prefer not to answer (6)

```{r}
ces2019 <- ces2019 %>%
  mutate(blame = case_when(
    pes19_blame == 6 ~ NA_real_,  
    TRUE ~ (5 - pes19_blame) / 4  
  ))
```


* pes19_deserve2: The welfare state makes people less willing to look after themselves.

* Strongly disagree (1)
* Somewhat disagree (2)
* Neither agree nor disagree (3)
* Somewhat agree (4)
* Strongly agree (5)
* Don't know/ Prefer not to answer (6)


```{r}
ces2019 <- ces2019 %>%
  mutate(deserve2 = case_when(
    pes19_deserve2 == 6 ~ NA_real_,
    TRUE ~ (5 - pes19_deserve2) / 4
  ))
```

* pes19_deserve1: If people really want to work, they can find a job.		
* Strongly disagree (1)
* Somewhat disagree (2)
* Neither agree nor disagree (3)
* Somewhat agree (4)
* Strongly agree (5)
* Don't know/ Prefer not to answer (6)

```{r}
ces2019 <- ces2019 %>%
  mutate(deserve1 = case_when(
    pes19_deserve1 == 6 ~ NA_real_, 
    TRUE ~ (5 - pes19_deserve1) / 4  
  ))
```

```{r}
ces2019 <- ces2019 %>%
    mutate(gap = NA,
           gap = ifelse(pes19_gap == 1, 0, gap),
           gap = ifelse(pes19_gap == 2, 0.25, gap),
           gap = ifelse(pes19_gap == 3, 0.5, gap),
           gap = ifelse(pes19_gap == 4, 0.75, gap),
           gap = ifelse(pes19_gap == 5, 1, gap))
```




## anti-immigration index

```{r}
ces2019 <- ces2019 %>%
   mutate(culture_harm = ifelse(pes19_nativism4 == 1, 0, NA), 
          culture_harm = ifelse(pes19_nativism4 == 2, 0.25, culture_harm),
          culture_harm = ifelse(pes19_nativism4 == 3, 0.50, culture_harm), 
          culture_harm = ifelse(pes19_nativism4 == 4, 0.75, culture_harm), 
          culture_harm = ifelse(pes19_nativism4 == 5, 1, culture_harm)) 

ces2019 <- ces2019 %>%
   mutate(imm_crim = ifelse(pes19_nativism5 == 1, 0, NA), 
          imm_crim = ifelse(pes19_nativism5 == 2, 0.25, imm_crim),
          imm_crim = ifelse(pes19_nativism5 == 3, 0.50, imm_crim), 
          imm_crim = ifelse(pes19_nativism5 == 4, 0.75, imm_crim), 
          imm_crim = ifelse(pes19_nativism5 == 5, 1, imm_crim)) 

ces2019 <- ces2019 %>%
   mutate(spend_imm_min = ifelse(cps19_spend_imm_min == 1, 1, NA), 
          spend_imm_min = ifelse(cps19_spend_imm_min == 2, 0.5, spend_imm_min),
          spend_imm_min = ifelse(cps19_spend_imm_min == 3, 0, spend_imm_min), 
          spend_imm_min = ifelse(cps19_spend_imm_min == 4, NA, spend_imm_min))


ces2019 <- ces2019 %>%
   mutate(admit_imm = ifelse(cps19_imm == 1, 0, NA), 
          admit_imm = ifelse(cps19_imm == 2, 1, admit_imm),
          admit_imm = ifelse(cps19_imm == 3, 0.5, admit_imm), 
          admit_imm = ifelse(cps19_imm == 4, NA, admit_imm))

ces2019 <- ces2019 %>%
    mutate(immig = (culture_harm + imm_crim + spend_imm_min + admit_imm) / 4)
```


## Vote Choice

```{r}
ces2019 <- ces2019 %>%
    mutate(vote_2019 = pes19_votechoice2019,
           vote_2019 = ifelse(is.na(vote_2019), cps19_v_advance, vote_2019))
```

```{r}
ces2019 <- ces2019 %>%
    mutate(pes19_turnout2019 = ifelse(is.na(pes19_turnout2019), pes19_turnout2019_v2, pes19_turnout2019))
```

```{r}
ces2019 <- ces2019 %>%
    mutate(vote_2019 = ifelse(pes19_turnout2019 > 1, 96, vote_2019)) %>%
    filter(pes19_turnout2019 != 3) %>%
    filter(pes19_turnout2019 != 8)
```

```{r}
ces2019 <- ces2019 %>%
    mutate(votePLC_19 = ifelse(vote_2019 == 1, 1, 0),
           voteCONS_19 = ifelse(vote_2019 == 2, 1, 0),
           voteNDP_19 = ifelse(vote_2019 == 3, 1, 0),
           voteBLOC_19  = ifelse(vote_2019 == 4, 1, 0))
```

```{r}
ces2019 <- ces2019 %>%
    mutate(vote_19_factor = NA,
           vote_19_factor = ifelse(vote_2019 == 1, "Liberal", vote_19_factor),
           vote_19_factor = ifelse(vote_2019 == 2, "Conservative", vote_19_factor),
           vote_19_factor = ifelse(vote_2019 == 3, "NDP", vote_19_factor),
           vote_19_factor = ifelse(vote_2019 == 4, "Bloc Quebecois", vote_19_factor),
           vote_19_factor = ifelse(vote_2019 == 5, "Green Party", vote_19_factor),
           vote_19_factor = ifelse(vote_2019 == 6, "People's Party", vote_19_factor),
#           vote_19_factor = ifelse(vote_2019 == 96, "Abstainer/DK", vote_19_factor),
           vote_19_factor = ifelse(vote_2019 == 96, NA, vote_19_factor))
```

```{r}
ces2019 <- ces2019 %>%
    mutate(vote_19_factor_abst = NA,
           vote_19_factor_abst = ifelse(vote_2019 == 1, "Liberal", vote_19_factor_abst),
           vote_19_factor_abst = ifelse(vote_2019 == 2, "Conservative", vote_19_factor_abst),
           vote_19_factor_abst = ifelse(vote_2019 == 3, "NDP", vote_19_factor_abst),
           vote_19_factor_abst = ifelse(vote_2019 == 4, "Bloc Quebecois", vote_19_factor_abst),
           vote_19_factor_abst = ifelse(vote_2019 == 5, "Green Party", vote_19_factor_abst),
           vote_19_factor_abst = ifelse(vote_2019 == 6, "People's Party", vote_19_factor_abst),
           vote_19_factor_abst = ifelse(vote_2019 == 96, "Abstainer/DK", vote_19_factor_abst))
```

```{r}
ces2019 <- ces2019 %>%
    mutate(right_bloc = ifelse(vote_19_factor == "Conservative" | vote_19_factor == "People's Party", 1, 0),
           right_bloc = ifelse(vote_19_factor == "Abstainer/DK", NA, right_bloc))
```

```{r}
ces2019 <- ces2019 %>%
    mutate(left_bloc = ifelse(vote_19_factor == "Green Party" | vote_19_factor == "Liberal" | vote_19_factor == "NDP", 1, 0),
           left_bloc = ifelse(vote_19_factor == "Abstainer/DK", NA, left_bloc))
```

```{r}
ces2019 <- ces2019 %>%
    mutate(left_bloc_abst = ifelse(vote_19_factor_abst == "Green Party" | vote_19_factor_abst == "Liberal" | vote_19_factor_abst == "NDP", 1, 0),
           left_bloc_abst = ifelse(vote_19_factor_abst == "Abstainer/DK", NA, left_bloc_abst))

ces2019 <- ces2019 %>%
    mutate(right_bloc_abst = ifelse(vote_19_factor_abst == "Conservative" | vote_19_factor_abst == "People's Party", 1, 0),
           right_bloc_abst = ifelse(vote_19_factor_abst == "Abstainer/DK", NA, right_bloc_abst))
```






## Post-transformation fine-tuning

```{r}
ces2019 <- ces2019 %>%
    mutate(weights = cps19_weight_general_all) %>%
    mutate(weights1 = pes19_weight_general_all) %>%
    mutate(weights2 = cps19_weight_general_restricted) %>%
    mutate(weights3 = pes19_weight_general_restricted) %>%
    mutate(region = factor(region))
```

```{r}
ces2019$education <- factor(ces2019$education, levels = c("No Schooling/Completed High School", "Some CEGEP/Post-Secondary", "Completed CEGEP/Post-Secondary", "Some University", "Bachelor's Degree", "Master's Degree/Doctorate"))

ces2019 <- ces2019 %>%
    mutate(religious_freq = factor(religious_freq, levels = c("Never/Almost never", "Once a month or more")))
```


```{r}
# generate index
ces2019 <- ces2019 %>%
  mutate(welfare_state = (stdofliving + inequal + govt_act_ineq) / 3,
         deservingness = (deserve1 + deserve2 + blame) / 3)

ces2019 <- ces2019 %>%
    filter(!is.na(welfare_state) & !is.na(deservingness))
```


## Assessing Robustness

```{r}
# alpha for both indexes

welfare_state_items <- ces2019[, c("stdofliving", "inequal", "govt_act_ineq")]
welfare_state_items <- welfare_state_items %>% na.omit()

alpha(welfare_state_items)

deservingness_items <- ces2019[, c("deserve1", "deserve2", "blame")]
deservingness_items <- deservingness_items %>% na.omit()

alpha(deservingness_items)
```

### Assessing the robustness of the anti-immigrant/nativism index

```{r}
immig_data <- ces2019 %>%
  select(culture_harm, imm_crim, spend_imm_min, admit_imm)

# Calculate Cronbach's alpha
cronbach_alpha <- alpha(immig_data)

# Display the result
print(cronbach_alpha)

df_transformed <- ces2019 %>%
  dplyr::select("blame", "deserve1", "deserve2", "stdofliving", "inequal", "govt_act_ineq") %>% na.omit()

 fa_result <- principal(df_transformed, nfactors=2, rotate="varimax")
 
 # View the results
print(fa_result)
 
 # Extract the factor loadings
factor_loadings <- fa_result$loadings
```


```{r}
 fa_result <- principal(df_transformed, nfactors=2, rotate="varimax")
 
 # View the results
print(fa_result)
 
 # Extract the factor loadings
factor_loadings <- fa_result$loadings
```


```{r}

# Convert the loadings to a matrix and then to a dataframe
loadings_matrix <- matrix(unlist(fa_result$loadings), ncol = 2)
rownames(loadings_matrix) <- rownames(fa_result$loadings)
loadings_df <- as.data.frame(loadings_matrix)
loadings_df$Variable <- rownames(loadings_df)
loadings_df <- melt(loadings_df, id.vars = "Variable", variable.name = "Factor", value.name = "Loading")
```

```{r}
# Create a named vector for descriptive variable labels
variable_labels <- c(inequal = "Income Inequality Problem",
                     deserve1 = "Job Availability Belief",
                     deserve2 = "Welfare State Dependency",
                     stdofliving = "Government Role in Standard of Living",
                     govt_act_ineq = "Government Action on Income Inequality",
                     blame = "Self-Blame for Not Getting Ahead")

# Update loadings_df with descriptive variable names
loadings_df$Variable <- factor(loadings_df$Variable, levels = names(variable_labels), labels = variable_labels)

# Update factor names to be more descriptive
factor_names <- c("V1" = "Deservingness", "V2" = "Welfare State")
loadings_df$Factor <- factor(loadings_df$Factor, levels = names(factor_names), labels = factor_names)
```


```{r}
library(dplyr)
library(ggplot2)

# Define the desired order of variables
desired_order <- c(
  "Government Action on Income Inequality",
  "Income Inequality Problem",
  "Government Role in Standard of Living",
  "Self-Blame for Not Getting Ahead",
  "Welfare State Dependency",
  "Job Availability Belief"
)

# Set the Variable column to a factor with the desired order
loadings_df <- loadings_df %>%
  mutate(Variable = factor(Variable, levels = desired_order))

# Plot with the specified order of variables
ggplot(loadings_df, aes(x = Variable, y = Loading, fill = Factor)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  coord_flip() +
  labs(x = "Attitudes Toward Redistribution", y = "Factor Loading") +
  scale_fill_manual(values = c("Welfare State" = "#088ec8", "Deservingness" = "#0d940d")) +
  theme_bw(base_size = 14) +
  theme(
    legend.title = element_blank(),
    legend.position = "bottom", # Remove the legend title
    axis.title.x = element_text(face = "bold"), # Bold X axis title
    axis.title.y = element_text(face = "bold") # Bold Y axis title
  )

# Save the plot
ggsave("factor_loadings_plot.png", width = 10, height = 5, dpi = 300)


```


```{r}
ggplot(loadings_df, aes(x = Variable, y = Loading, fill = Factor)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  coord_flip() +
  labs(x = "Attitudes Toward Redistribution", y = "Factor Loading") +
  scale_fill_manual(values = c("Welfare State" = "#088ec8", "Deservingness" = "#0d940d")) +
  theme_bw(base_size = 14) +
  theme(
    legend.title = element_blank(),
    legend.position = "bottom", # Remove the legend title
    axis.title.x = element_text(face = "bold"), # Bold X axis title
    axis.title.y = element_text(face = "bold") # Bold Y axis title
  )

# Save the plot
ggsave("factor_loadings_plot.png", width = 10, height = 5, dpi = 300)


```

```{r}
loadings_df <- loadings_df %>%
  spread(key = Factor, value = Loading)
```


```{r}
loadings_df <- loadings_df %>%
  mutate(HighestLoading = ifelse(`Welfare State` > `Deservingness`, "Welfare State", "Deservingness")) %>%
  gather(key = "Factor", value = "Loading", -Variable, -HighestLoading)
```


```{r}
loadings_df$ColorFill <- ifelse(loadings_df$Factor == loadings_df$HighestLoading, loadings_df$Factor, "Other")

library(ggplot2)

library(ggplot2)

ggplot(loadings_df, aes(x = Variable, y = Loading, fill = ColorFill)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_manual(values = c("Welfare State" = "#088ec8", 
                               "Deservingness" = "#0d940d", 
                               "Other" = "grey"),
                    breaks = c("Welfare State", "Deservingness")) + # Specify legend items
  coord_flip() +
  labs(x = "Attitudes Toward Redistribution", y = "Factor Loading") +
  theme_bw(base_size = 14) +
  theme(
    legend.title = element_blank(),
    legend.position = "bottom",
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold")
  )

ggsave("factor_loadings_plot_precise.png", width = 10, height = 5, dpi = 300)
```




```{r}
df <- ces2019 %>%
    dplyr::select(c(vote_19_factor, vote_19_factor_abst, right_bloc, left_bloc, right_bloc_abst, left_bloc_abst, education, agegroup, gender, religious_freq, language, minovis, urban_rural, deservingness, welfare_state, weights, weights1, weights2, weights3, region, immig))
```


```{r}
# generate dichotomous variable for each party
df <- df %>%
    mutate(bloc = ifelse(vote_19_factor == "Bloc Quebecois", 1, 0),
           conservative = ifelse(vote_19_factor == "Conservative", 1, 0),
           green = ifelse(vote_19_factor == "Green Party", 1, 0),
           liberal = ifelse(vote_19_factor == "Liberal", 1, 0),
           ndp = ifelse(vote_19_factor == "NDP", 1, 0),
           people_party = ifelse(vote_19_factor == "People's Party", 1, 0))
```

```{r}
# generate dichotomous variable for each party
df <- df %>%
    mutate(bloc_abst = ifelse(vote_19_factor_abst == "Bloc Quebecois", 1, 0),
           conservative_abst = ifelse(vote_19_factor_abst == "Conservative", 1, 0),
           green_abst = ifelse(vote_19_factor_abst == "Green Party", 1, 0),
           liberal_abst = ifelse(vote_19_factor_abst == "Liberal", 1, 0),
           ndp_abst = ifelse(vote_19_factor_abst == "NDP", 1, 0),
           people_party_abst = ifelse(vote_19_factor_abst == "People's Party", 1, 0),
           abst = ifelse(vote_19_factor_abst == "Abstainer/DK", 1, 0))
```


## Descriptive Table

```{r}
df <- df %>%
    mutate(gender = ifelse(gender == 1, "Woman", "Man"),
           minovis = ifelse(minovis == 1, "Visible Minority", "Other"))

datasummary(vote_19_factor + education + agegroup +  gender +  religious_freq + minovis + language + urban_rural + region  ~ mean + Percent() + N, data = df, output = "descriptive_table.docx")

save(df, file = "ces2019_analyses.RData")
```



```{r}
# subset for STATA analysis
df <- ces2019 %>%
    filter(!is.na(vote_19_factor_abst) & !is.na(education) & !is.na(agegroup) & !is.na(gender) & !is.na(religious_freq) & !is.na(language) & !is.na(minovis) & !is.na(urban_rural) & !is.na(deservingness) & !is.na(welfare_state))

write_dta(df, "ces2019_analyses.dta")
```

